### 2020-06-25 manuscript/manuscript.rmd
made pass through Variation in results across published analyses

analyses_published_processed = 
  analyses_published %>%
  dplyr::filter( !grepl( 'Recoding', model_combined ) ) %>%
  dplyr::rename( model = model_combined ) %>%
  select( model, clade, final, inference ) %>%
  dplyr::rename( result = final ) %>%
  mutate( status = "published" )

analyses_new_processed = 
  analyses_new %>%
  mutate(model = factor(model)) %>%
  select( model, clade, result, inference ) %>%
  mutate( status = "new" )

@@

**Fig. 3.**  A summary of phylogenomic results from the previous studies and reanalysis conducted in this study. Each shape represents an phylogenetic analysis that has been analyzed by ML (designated by the circle) or BI inference (designated by the triangle). Analyses with a support of bootstrap values less than 90% or a posterior probability less than 0.95 are considered as unresolved. (A) Analyses transcribed from the literature (related to Supplementary Table XX). (B) New phylogenomic analyses were conducted in this study (related to Supplementary Table XX).

We parsed `r nrow(analyses_published)` previous phylogenetic analyses from `r length(unique(analyses_published$manuscript))` studies (Fig. 3A and Supplementary Table 2). The conclusions of five studies strongly favor of Porifera-sister and 10 favor Ctenophora-sister (Extended Data Table 1). Three studies are based entirely on previously published data, and the remainder add data for one or more species.

All published analyses that use models with site-homogeneous equilibrium frequencies (*i.e.*, models that do not include CAT) strongly support Ctenophora-sister, regardless of other factors such as outgroup choice (Fig. 3A). Some analyses that exclude more distant outgroups and use models with site-heterogeneious CAT equilibrium frequencies support Porifera-sister. Because CAT is so parameter rich, many analyses that use a CAT model do not converge, and some of the unconvered analyses have an unresolved animal root. Although model adequacy assessments generally favor GTR+CAT over Poisson+CAT or site-homogeneous models [@pisani2015genomic; @feuda2017improved], Porifera-sister is not always supported since several analyses strongly in favor Ctenophora-sister even with only choanofalgellates are used as outgroups (Fig. 3A). Thus, our summary of previous phylogenetic analyses strongly indicated that the trend of the support of Porifera-sister is generally increasing with the use of more complex CAT models and more closely related outgroups in several representative matrices, whereas Ctenophora-sister is supported in most other conditions.

Inclusion of more distant related (e.g. Fungi) outgroup taxa tend to recover Ctenophora-sister, whereas increasing support of Porifera-sister often recovered when more closely related outgroup taxa were used with site-heterogeneous Poisson-CAT models [@pisani2015genomic].

@@

Here we reanalyzed 17 matrices from 13 studies that are used as main conclusions from their original paper (Extended Data Table 1). We also progressively trimmed each data matrix with three level of outgroup choice if applicable (Fig 1C). We first tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C60 models using ModelFinder [@kalyaanamoorthy2017modelfinder] in IQ-TREE (except only LG+C60 models are used in 3 large matrices). We found in all cases, WAG/LG+C60 models fit these matrices better than the site-homogeneous models under BIC criteria and we then inferred support under the best-fit model from IQ-TREE (Extended Data Table 1). We then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG, GTR and Poisson+C60. Moreover, we also used the best-fit model identified above with the removal of all distantly related outgroups and only include Choanoflagellata as outgroups. Interestingly, with the exception of Moroz2014_3d and all Nosenko2013 matrices that supports neither Porifera-sister or Ctenophora-sister, all analysis conducted herein in IQ-TREE strongly supported the ctenophore-sister hypothesis, even with CAT-like, site-heterogeneous C60 models (Poisson+C60 and WAG/LG+C60) with the most closely related outgroups (Fig. 3B; Supplementary Table 3).

>>>>

analyses_published_processed = 
  analyses_published %>%
  dplyr::filter( !grepl( 'Recoding', model_combined ) ) %>%
  dplyr::rename( model = model_combined ) %>%
  select( model, clade, final, inference, manuscript, matrix ) %>%
  dplyr::rename( result = final ) %>%
  mutate( status = "published" )

analyses_new_processed = 
  analyses_new %>%
  mutate(model = factor(model)) %>%
  select( model, clade, result, inference, matrix ) %>%
  mutate( manuscript = NA ) %>%
  mutate( status = "new" )

@@

**Fig. 3.**  A summary of phylogenomic results from the previous studies and reanalysis conducted in this study. Each point represents a phylogenetic Analyses with a support of bootstrap values less than `r bootstrap_threshold`% or a posterior probability less than `r posterior_prob_threshold`% are considered as unresolved. The clades are orgainized with increased outgroup sampling higher on the vertical axis, and the models are organized in general by increasing complexity to the right on the horizontal axis. (A) Analyses transcribed from the literature (related to Supplementary Table XX). (B) New phylogenomic analyses conducted in this study (related to Supplementary Table XX).

We parsed `r analyses_processed %>% filter(status=="published") %>% nrow()` previous phylogenetic analyses from `r analyses_processed %>% filter(status=="published") %>% pull (manuscript) %>% unique() %>% length()` studies (Fig. 3A and Supplementary Table 2). The conclusions of five studies strongly favor of Porifera-sister and 10 favor Ctenophora-sister (Extended Data Table 1). Three studies are based entirely on previously published data, and the remainder add data for one or more species.

As others have noted (XX), our summary of previous phylogenetic analyses (Fig. 3A) show that Ctenophora-sister is supported over analyses that span the full range of outgroup sampling and models used to date, including some analyses with restricted outgroup sampling and models that accommondate site-heterogeneous equilibrium frequencies with CAT. The only analyses that support Porifera-sister ahveh reduced outgroup sampling (Choanimalia, Holozoa) and site heterogeneous models with CAT. Model adequacy assessments generally favor GTR+CAT over Poisson+CAT or site-homogeneous models [@pisani2015genomic; @feuda2017improved], but because CAT is so parameter rich, many analyses that use a model with CAT do not converge. Some of the unconvered analyses have an unresolved animal root. The fact that Porifera-sister is recovered only for particular models with particular outgroup sampling indicates that model and outgroup interact, and that this interaction is fundamental to understanding the range of results obtained across analyses.

@@

Here we reanalyzed 17 matrices from 13 studies that are used as main conclusions from their original paper (Extended Data Table 1). We also progressively trimmed each data matrix with three level of outgroup choice if applicable, original outgroup sampling permitting (Fig 1C). We first tested a variety of models for each matrix and compared the relative fit of site-homogeneous and site-heterogeneous C60 models using ModelFinder [@kalyaanamoorthy2017modelfinder] in IQ-TREE (except only LG+C60 models are used in 3 large matrices). We found in all cases, WAG/LG+C60 models fit these matrices better than the site-homogeneous models under BIC criteria and we then inferred support under the best-fit model from IQ-TREE (Extended Data Table 1). We then analyzed every matrix under a panel of standard site-heterogeneous and site-homogeneous models, including WAG, GTR and Poisson+C60. Moreover, we also used the best-fit model identified above with the removal of all distantly related outgroups and only include Choanoflagellata as outgroups. Interestingly, with the exception of Moroz2014_3d and all Nosenko2013 matrices that supports neither Porifera-sister or Ctenophora-sister, all analysis conducted herein in IQ-TREE strongly supported the ctenophore-sister hypothesis, even with CAT-like, site-heterogeneous C60 models (Poisson+C60 and WAG/LG+C60) with the most closely related outgroups (Fig. 3B; Supplementary Table 3).
